Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-19 PAPER.tex
made some adrian edits until section 3.1.3

By about one billion years after the Big Bang ($z \sim 6$), the first stars and galaxies are thought to have ionized all the 
neutral hydrogen that dominated the baryonic matter content in the Universe. This transition period, during which the first 
luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas 
into a plasma, is known as the Epoch of Reionization (EoR). The EoR is a relatively unexplored era in our \textit{cosmic dawn}, which spans from the birth of the first stars to the full reionization of the intergalactic medium (IGM). Its 
history encodes important information regarding the nature of the first galaxies and the processes of structure formation. 
Direct measurements of the EoR would unlock powerful characteristics about the IGM, revealing connections 
between the matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured 
web of galaxies we observe today \citep[for a review, see][]{furlanetto_et_al2006}.

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned.

@@

We begin with our data vector, $\textbf{x}$, which contains our measured visibilities for a single baseline in Jy. It has length $N_{t} \times N_{f}$, but in practice we manipulate it as an array with dimensions ($N_{t}, N_{f}$), where 
$N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of 
the Fourier transform of the sky along $2$ spatial dimensions (using the flat-sky approximation), and since we are interested in $3$-dimensional Fourier-modes 
we only need to take one Fourier transform of our visibilities along the line-of-sight dimension. We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\widehat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}
Here, \noindent $\textbf{Q}$ is a family of matrices that takes our frequency-domain visibilities and Fourier-transforms them into 
power spectrum space, while also converting from Jy to 
Kelvin and taking into account cosmological scalings. It is evaluated as $\textbf{Q}^{\alpha} \equiv \frac{\delta\textbf{C}}{\delta p_{\alpha}}$, or the derivative of the covariance with respect to the true bandpower in $\alpha$, where $\alpha$ denotes a waveband in $k_{\parallel}$ ($k_{\parallel}$, a cosmological wavenumber, is the 
Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b})). $\textbf{R}$ is a weighting matrix --- as 
an example, inverse covariance weighting would set $\textbf{R} = \textbf{C}^{-1}$ and an unweighted case would use $
\textbf{R} = \textbf{I}$, the identity matrix.

@@

In the next three sections, we investigate the effects of weighting matrices on signal loss by experimenting with different 
matrices $\textbf{R}$ and examine their impact on the resulting power spectra estimate $\widehat{\textbf{p}}$. Our goal in 
experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we 
purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section 
\ref{sec:Sigloss}.

@@

\noindent assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix is therefore $
\textbf{R} = \widehat{\textbf{C}}^{-1}$.

@@

If $\textbf{C}$ is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a 
multiplicative bias (per $k$) to estimates of the signal. For a mathematical derivation of signal loss arising from a data-estimated covariance matrix, see Appendix \ref{sec:sigloss_appendix}. Here we will describe the origin of this signal loss 
intuitively.

@@

In general, the strongest modes of $\widehat{\textbf{C}}$ (highest eigenvalues) are expected to be dominated by bright foregrounds 
--- the most prominent shapes in the data. For these `strong' modes, where foregrounds outshine the EoR signal, $
\widehat{\textbf{C}} \sim \textbf{C}$ and down-weighting these modes is beneficial. This is demonstrated in the toy model by the 
successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid shape and 
assigns it the highest eigenvalue (the first peak in Figure \ref{fig:toy_sigloss2}). In Figure \ref{fig:toy_sigloss12} we show the 
covariances of our toy model datasets along with inverse covariance weighted data. The foreground sinusoid is clearly visible in 
$\widehat{\textbf{C}}_{FG}$ but effectively absent from $\widehat{\textbf{C}}^{-1}_{FG}\textbf{x}_{FG}$.

@@

The danger of an empirically estimated covariance matrix comes from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the empirical covariance believes these modes should be down-weighted more than the true covariance would. For these modes, the error that results from not computing $\textbf{C}$ accurately (due to a finite data sample), has a larger impact on the `weak' modes than the `strong' modes. In other words, mis-estimating $\textbf{C}$ affects EoR-dominated eigenmodes more than FG-dominated modes, and it is primarily the error associated in down-weighting these `weak' modes that leads to signal loss.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield minimal 
signal loss. Recall that our toy model foreground is a sinusoid, so it can be perfectly described by a single eigenmode. Using the full dataset's (foreground plus EoR signal) empirical covariance, we can 
project out the first eigenmode (i.e. use the true eigenvalue of the first eigenmode, but flatten the rest of the spectrum to have eigenvalues of 1), thereby down-weighting the foreground-dominated mode more than the rest of the modes. Hence, we are not using the relative weighting information from all the weaker eigenmodes which have comparatively larger errors. Altering $\widehat{\textbf{C}}$ as such is one example of a regularization method, in which we are 
changing $\widehat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more similar to that of $\textbf{I}$. The resulting 
power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, 
demonstrating that if we can disentangle the foreground-dominating modes from EoR-dominating modes, we can down-weight 
them without signal loss. There are several other ways to regularize $\widehat{\textbf{C}}$, and we will discuss some in Section 
\ref{sec:otherweight}.

@@

We have shown how signal loss can arise due to inaccurately characterizing weak eigenmodes (EoR dominated modes) with a 
data-estimated covariance. We will next show how this effect is exaggerated by reducing the total number of independent 
samples in a dataset.

A fringe-rate filter is an analysis technique designed to maximize sensitivity by integrating in time (\citealt{parsons_et_al2016}). 
Rather than a traditional box-car average, a time domain filter can be designed to up-weight temporal modes consistent with 
the sidereal motion on the sky, while down-weighting modes which are noise-like.

@@

With 
fringe-rate filtering resulting in fewer independent modes, it becomes more difficult for the empirical covariance to estimate the 
true covariance matrix. This can be quantified by evaluating a convergence metric $\varepsilon(\Chat)$ for the empirical covariance, which we define as
\begin{equation}
\label{eq:converge}
\varepsilon (\Chat) \equiv \sqrt{\frac{\sum_{ij} (\widehat{C}_{ij} - {C}_{ij})^2}{\sum_{ij} {C}_{ij}^2}},
\end{equation}
where $\C$ is true covariance matrix. In Figure \ref{fig:toy_sigloss16} we show this convergence statistic as a function of the number of independent ensemble realizations in one's simulations (horizontal axis) and the number of independent samples in the data following fringe-rate filtering (different curve colors). With more independent time samples in the data, one converges to the true covariance more quickly.

@@

Our second major $21$ cm power spectrum theme is error estimation, as we desire robust methods for determining accurate 
confidence intervals for our measurements. Two popular ways of calculating errors on a power spectrum 
measurement are calculating the variance of a dataset, and computing a theoretical error estimate based on an instrument's 
system temperature and observational parameters. In a perfect world, both methods would match up. However, in practice the 
two don't always agree due to a number of factors, including time, frequency, and antenna-dependent noise and non-uniform 
weightings.

For PAPER's analysis, we choose to compute error bars on our measurements that have been derived from its inherent 
variance. A common technique used to do this is bootstrapping. For pedagogical purposes, we first define the technique of 
bootstrapping and then illustrate one of its pitfalls through a toy model.

Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, measurements (like power 
spectra) can be made from different samples of data. Each of these measurements is a different realization drawn from some underlying distribution, and realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common 
between them. Hence, through the process of resampling and averaging along different axes, we can estimate error bars for 
our results which represent the underlying distribution of values that are allowed by our measurements (\citealt{efron_tibshirani1994}; \citealt{andrae2010}).

Suppose we have $N$ different measurements targeting the same quantity ($N$ power spectrum measurements). 
Bootstrapping means that we form $N_{boot}$ (often a large number) bootstraps, where each bootstrap is a random selection 
of the $N$ measurements. Bootstraps each have dimensions of $N$, and the values populated into each bootstrap are drawn 
from the original set of measurements with replacement (i.e. every $N$th slot is filled randomly for each bootstrap). Next we take 
the mean of each bootstrap to collapse it from an array of length $N$ to a single number (we are interested in the mean statistic 
here, but any function of interest can be applied to each bootstrap as long as it's the same function for each one). The error (on 
the mean) is then computed as the standard deviation across all bootstraps.

We must be careful in distinguishing $N_{boot}$, the number of bootstraps, from $N$, the number of samples, or elements, or 
values, that comprise a bootstrap. In the toy models presented in this section, $N_{boot}$ is typically large, and the standard 
deviation across bootstraps (the error we are computing) converges for large $N_{boot}$. It is the value of $N$ that we must be 
careful in setting, as described in the rest of this section.

@@

This example highlights the importance of understanding how analysis techniques (e.g. fringe-rate filtering) can affect a 
common statistical procedure like bootstrapping. Bootstrapping as a means of estimating power spectrum errors from real 
fringe-rate filtered data requires knowledge of the number of independent samples, which is not always a trivial task. For 
example, computing the effective number of independent samples of fringe-rate filtered data is not as simple as counting the 
number of averages performed. Down-sampling a time-averaged signal is straightforward using a boxcar average, but non-trivial with a more complicated convolution function that has long tails. Hence, we do not recommend bootstrapping unless the 
number of independent samples along the axis that is being re-sampled is well-determined.

@@

Foreground bias is perhaps one of the main factors limiting $21$ cm results. Foreground signals lie $\sim4$-$5$ orders of 
magnitude above the cosmological signal. Though there are many techniques proposed for removing foregrounds (see e.g. \citealt{vedantham_et_al2012}; \citealt{parsons_et_al2012a}; \citealt{parsons_et_al2012b}; \citealt{dillon_et_al2013a}; \citealt{wang_et_al2013}; \citealt{parsons_et_al2014}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2015}; \citealt{pober_et_al2016a}; \citealt{trott_et_al2016}), most 
experiments currently remain limited by residuals rather than noise, especially at low $k$.

@@

Although smooth spectrum foregrounds preferentially show up at low delay, or low $k$ modes, their isolation within the wedge is not perfect.  In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond 
the length of a baseline are often still contaminated at a low level. This leakage, particularly at low $k$'s, can be attributed to 
convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum 
foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum, the signal chain, and the 
antenna response, all of which could smear out the foregrounds and cause leakage outside the wedge \citep[e.g.][]{ewall-wice_et_al2017}.

There are analysis techniques to mitigate the effects of foreground leakage and prevent information from low $k's$ from 
spreading to high $k$ values. For example, narrow window functions in delay-space can be used to minimize the covariance of a particular 
$k$ value with other ones (\citealt{liu_et_al2014b}). In other words, one can construct an estimator using OQE that forces a 
window function to have a minimum response to low $k$ values. The window function used in A15 is constructed in such a way, 
specifically to prevent foregrounds that live at low $k's$ from contaminating higher $k$-modes (see Section \ref{sec:Bias}). 
Minimizing foreground leakage in this way however, comes with the trade-off of compromising power spectrum sensitivity, since 
narrow window functions increases errors for each $k$-mode (\citealt{liu_et_al2014b}).

Determining the source of detections at higher $k$'s is more difficult. In previous power spectrum results, these detections have been explained as instrumental systematics, particularly time-variable cross talk, RFI, cable reflections, and calibration errors (A15; \citealt{parsons_et_al2014}; \citealt{dillon_et_al2014}; \citealt{beardsley_et_al2016}; \citealt{patil_et_al2017}). In the next section (Section \ref{sec:JackknifeOverview}), we will present some tests that can help distinguish these excesses from that of EoR.

@@

Another type of noise bias can stem from the spurious cross-coupling of signals between antennas. This excess is known as 
instrumental crosstalk and is an inadvertent correlation between two independent measurements via a coupled signal path. 
Crosstalk appears as a constant phase bias in time in visibilities, and it varies slowly compared to the typical fringe-rates of 
sources. Because it is slow-varying, crosstalk can be suppressed using time-averages or fringe-rate filters. However, there 
remains a possibility that power spectrum detections are caused by residual, low-level crosstalk which survived any 
suppression techniques.

@@

To mimic the presence of a systematic in part of the measurement, we add a toy sinusoid foreground, similar to the one used in 
Section \ref{sec:toymodel}, to the first $100$ time integrations of both measurements. This represents a foreground signal 
present in, for example, the first half of the LST range used for analysis, but not the second half. Mathematically, if  $\textbf{n}$ 
is noise, $\textbf{e}$ is the EoR signal, and $\textbf{fg}$ is the foreground signal, the two measurements (which are cross-multiplied to form power spectra) can be written as:

\begin{equation}
\label{eq:bias1}
\textbf{x}_{a} = \textbf{n}_{a} + \textbf{e} + \textbf{fg}
\end{equation}

@@

\begin{equation}
\label{eq:bias2}
\textbf{x}_{b} = \textbf{n}_{b} + \textbf{e} + \textbf{fg}.
\end{equation}

@@

\begin{eqnarray}
\textbf{x}_{1} &=& \textbf{n} + \textbf{e} + \textbf{fg} \\
\textbf{x}_{2} &=& \textbf{n} + \textbf{e} 
\end{eqnarray}

@@

We form three different power spectrum estimates shown in Figure \ref{fig:toy_bias1}. The first is a null test where we subtract $
\textbf{x}_{2}$ from $\textbf{x}_{1}$ for both measurements $a$ and $b$. This is equivalent to splitting up a full dataset along 
an axis (in this case, time) and subtracting the two to remove sky signal that should ideally be present in both. We cross-
multiply measurements $a$ and $b$ to form an un-biased (thermal noise-wise) estimate (blue curve). The second estimate, 
shown in red, is the same null test with the foreground systematic removed (eliminate $\textbf{fg}$ in Equations \ref{eq:bias1} 
and \ref{eq:bias2}). Finally, we also show the noise power spectrum (green).

@@

\subsubsection{Signal Loss Methodology} 
\label{sec:siglossmethod}
To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method as in A15, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.\footnote{One 
specific change from A15 is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss, %(by a factor of $\sim$$10$), 
likely due to the use of the fringe-rate filter as a simulator. However, this pipeline difference, while significant, is not the dominant reason why signal loss was under-estimated in A15 (the dominant reason is explained in the main text).}

Suppose that $\textbf{e}$ is the injected EoR (at some amplitude level), and $\textbf{x}$ is our data. We define $\textbf{r}
$ to be the data plus the EoR signal:

>>>>

By about one billion years after the Big Bang ($z \sim 6$), the first stars and galaxies are thought to have ionized all the 
neutral hydrogen that dominated the baryonic matter content in the Universe. This transition period, during which the first 
luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas 
into a plasma, is known as the Epoch of Reionization (EoR). The EoR is a relatively unexplored era in our \textit{cosmic dawn}, which spans from the birth of the first stars to the full reionization of the intergalactic medium (IGM). Its 
history encodes important information regarding the nature of the first galaxies and the processes of structure formation. 
Direct measurements of the EoR would unlock powerful characteristics about the IGM, revealing connections 
between the matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured 
web of galaxies we observe today (for a review, see \citet{furlanetto_et_al2006} and \citet{loeb_furlanetto_2013}).

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, errors, and bias) that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned.

@@

We begin with our data vector, $\textbf{x}$, which contains our measured visibilities for a single baseline in Jy. It has length $N_{t} \times N_{f}$, but in practice we manipulate it as an array with dimensions ($N_{f}, N_{t}$), where 
$N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of 
the Fourier transform of the sky along $2$ spatial dimensions (using the flat-sky approximation), and since we are interested in $3$-dimensional Fourier-modes 
we only need to take one Fourier transform of our visibilities along the line-of-sight dimension. We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\widehat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}
Here, \noindent $\textbf{Q}$ is a family of matrices that takes our frequency-domain visibilities and Fourier transforms them into 
power spectrum space, while also converting from Jy to 
Kelvin and taking into account cosmological scalings. It is evaluated as $\textbf{Q}^{\alpha} \equiv \frac{\partial\textbf{C}}{\partial p_{\alpha}}$ (which has dimensions ($N_{f}$,$N_{f}$)), or the derivative of the covariance with respect to the true bandpower in $\alpha$, where $\alpha$ denotes a waveband in $k_{\parallel}$ ($k_{\parallel}$, a cosmological wavenumber, is the 
Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}). $\textbf{R}$ is a weighting matrix with dimensions ($N_{f}$,$N_{f}$) --- as 
an example, inverse covariance weighting would set $\textbf{R} = \textbf{C}^{-1}$ and an unweighted case would use $
\textbf{R} = \textbf{I}$, the identity matrix.

@@

In the next three sections, we use toy models to investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examining their impact on the resulting power spectra estimate $\widehat{\textbf{p}}$. Our goal in 
experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we 
purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section 
\ref{sec:Sigloss}.

@@

\noindent assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix for inverse covariance weighting is then $
\textbf{R} = \widehat{\textbf{C}}^{-1}$.

@@

If $\textbf{C}$ is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a 
multiplicative bias (per $k$) to estimates of the signal. For a toy model mathematical derivation of signal loss arising from a data-estimated covariance matrix, see Appendix \ref{sec:sigloss_appendix}. Here we will describe the origin of this signal loss 
intuitively.

@@

In general, the strongest modes of $\widehat{\textbf{C}}$ (highest eigenvalues) are expected to be dominated by bright foregrounds 
--- the most prominent shapes in the data. For these `strong' modes, where foregrounds outshine the EoR signal, $
\widehat{\textbf{C}} \sim \textbf{C}$ and down-weighting these modes is beneficial. This is demonstrated in the toy model by the 
successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid shape and 
assigns it the highest eigenvalue (the first peak in Figure \ref{fig:toy_sigloss2}). In Figure \ref{fig:toy_sigloss12} we show the 
covariances of our toy model datasets along with inverse covariance weighted data. The foreground sinusoid is clearly visible in 
$\widehat{\textbf{C}}_{FG}$.% but effectively absent from $\widehat{\textbf{C}}^{-1}_{FG}\textbf{x}_{FG}$.

@@

The danger of an empirically estimated covariance matrix comes from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the empirical covariance believes these modes should be down-weighted more than the true covariance would. For these modes, the error that results from not computing $\textbf{C}$ accurately (due to a finite data sample), has a larger impact on the `weak' modes than the `strong' modes. In other words, mis-estimating $\textbf{C}$ affects EoR-dominated eigenmodes more than FG-dominated modes, and it is primarily the coupling between these mis-estimated modes to the data that leads to signal loss.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield minimal 
signal loss. Recall that our toy model foreground is a sinusoid, so it can be perfectly described by a single eigenmode. Using the full dataset's (foreground plus EoR signal) empirical covariance, we can 
project out the first eigenmode (i.e. use the true eigenvalue of the first eigenmode, but flatten the rest of the spectrum to have eigenvalues of 1), thereby down-weighting the foreground-dominated mode more than the rest of the modes. Hence, we are not using the relative weighting information from all the weaker eigenmodes which have comparatively larger errors. Altering $\widehat{\textbf{C}}$ as such is one specific example of a regularization method for this toy model, in which we are 
changing $\widehat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more similar to that of $\textbf{I}$. The resulting 
power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, 
demonstrating that if we can disentangle the foreground-dominating modes from EoR-dominating modes, we can down-weight 
them with negligible signal loss. There are several other ways to regularize $\widehat{\textbf{C}}$, and we will discuss some in Section 
\ref{sec:otherweight}.

@@

We have shown how signal loss can arise due to the coupling of weak eigenmodes (EoR dominated modes) to the data. We will next show how this effect is exaggerated by reducing the total number of independent 
samples in a dataset.

A fringe-rate filter is an analysis technique designed to maximize sensitivity by integrating in time (\citealt{parsons_et_al2016}). 
Rather than a traditional box-car average in time, a time domain filter can be designed to up-weight temporal modes consistent with 
the sidereal motion on the sky, while down-weighting modes which are noise-like.

@@

With 
fringe-rate filtering resulting in fewer independent modes, it becomes more difficult for the empirical covariance to estimate the 
true fringe-rate filtered case covariance matrix. This can be quantified by evaluating a convergence metric $\varepsilon(\Chat)$ for the empirical covariance, which we define as
\begin{equation}
\label{eq:converge}
\varepsilon (\Chat) \equiv \sqrt{\frac{\sum_{ij} (\widehat{C}_{ij} - {C}_{ij})^2}{\sum_{ij} {C}_{ij}^2}},
\end{equation}
where $\C$ is true covariance matrix. In Figure \ref{fig:toy_sigloss16} we show this convergence statistic as a function of the number of independent ensemble realizations in one's simulations (horizontal axis) and the number of independent samples in the data following fringe-rate filtering (different curve colors). With more independent time samples in the data, one converges to the true fringe-rate filtered covariance more quickly.

@@

Our second major $21$ cm power spectrum theme is error estimation, as we desire robust methods for determining accurate 
confidence intervals for our measurements. Two popular ways of calculating errors on a power spectrum 
measurement are calculating the variance of a dataset, and computing a theoretical error estimate based on an instrument's 
system temperature and observational parameters. In a perfect world, both methods would match up. However, in practice the 
two don't always agree due to a number of factors, including possible non-Gaussianities in the noise properties of our instruments and non-uniform 
weightings.

For PAPER's analysis, we choose to compute error bars that have been derived from the inherent 
variance of our measurements. A common technique used to do this is bootstrapping. For pedagogical purposes, we first define the technique of 
bootstrapping and then illustrate one of its pitfalls through a toy model.

Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, measurements (like power 
spectra) can be made from different samples of data. Each of these measurements is a different realization drawn from some underlying distribution, and realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common 
between them. Through the process of resampling and averaging along different axes, we can estimate error bars for 
our results which represent the underlying distribution of values that are allowed by our measurements (\citealt{efron_tibshirani1994}; \citealt{andrae2010}).

Suppose we have $N$ different measurements targeting the same quantity ($N$ power spectrum measurements made along different axes, such as baselines or times). 
Bootstrapping means that we form $N_{boot}$ (often a large number) bootstraps, where each bootstrap is a random selection 
of the $N$ measurements. Bootstraps each have dimensions of $N$, and the values populated into each bootstrap are drawn 
from the original set of measurements with replacement (i.e. every $n^{th}$ slot in $N$ is filled randomly for each bootstrap). Next we take 
the mean of each bootstrap to collapse it from an array of length $N$ to a single number (we are interested in the mean statistic 
here, but any function of interest can be applied to each bootstrap as long as it's the same function for each one). The error (on 
the mean) is then computed as the standard deviation across all bootstraps.

We must be careful in distinguishing $N_{boot}$, the number of bootstraps, from $N$, the number of samples, or elements, or 
values, that comprise a bootstrap. In the toy models presented in this section, $N_{boot}$ is typically large, and the standard 
deviation across bootstraps (the error we are computing) converges for large $N_{boot}$. Typically $N$ is a straightforward value to set that just depends on the experiment. However, we will illustrate one case in which it is not simply the number of samples along the axis that is being re-sampled. More specifically, we will see that $N$ depends on sample independence and may not always be straightforward to approximate.

@@

This example highlights the importance of understanding how analysis techniques (e.g. fringe-rate filtering) can affect a 
common statistical procedure like bootstrapping. Bootstrapping as a means of estimating power spectrum errors from real 
fringe-rate filtered data requires knowledge of the number of independent samples, which is not always a trivial task. For 
example, computing the effective number of independent samples of fringe-rate filtered data is not as simple as counting the 
number of averages performed. Down-sampling a time-averaged signal is straightforward using a boxcar average, but non-trivial with a more complicated convolution function that has long tails. Hence, we do not recommend bootstrapping unless the 
number of independent samples along the axis that is being re-sampled is well-determined. In Section \ref{sec:Boot}, we explain how our bootstrapping procedure has changed in PAPER's analysis in light of this issue.

@@

In Section \ref{sec:SiglossOverview}, we discussed signal loss as a form of multiplicative bias to estimates of the signal. Foregrounds are another type of bias, but an additive one instead of multiplicative. Foreground bias is perhaps one of the main factors limiting $21$ cm results, as foreground signals lie $\sim4$-$5$ orders of 
magnitude above the cosmological signal. Though there are many techniques proposed for removing foregrounds (see e.g. \citealt{vedantham_et_al2012}; \citealt{parsons_et_al2012a}; \citealt{parsons_et_al2012b}; \citealt{dillon_et_al2013a}; \citealt{wang_et_al2013}; \citealt{parsons_et_al2014}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2015}; \citealt{pober_et_al2016a}; \citealt{trott_et_al2016}), most 
experiments currently remain limited by residuals rather than noise, especially at low $k$.

@@

Although smooth spectrum foregrounds preferentially show up at low delay, or low $k$ modes, their isolation within the wedge is not perfect. In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond 
the delay associated with the length of a baseline are often still contaminated at a low level. This leakage, particularly at low $k$'s, can be attributed to 
convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum 
foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum, the signal chain, and the 
antenna response, all of which could smear out the foregrounds and cause leakage outside the wedge \citep[e.g.][]{ewall-wice_et_al2017}.

There are analysis techniques to mitigate the effects of foreground leakage and prevent information from low $k's$ from 
spreading to high $k$ values. For example, narrow window functions in delay-space can be used to minimize the leakage from a particular 
$k$ value into other ones (\citealt{liu_et_al2014b}). In other words, one can construct an estimator using OQE that forces a 
window function to have a minimum response to low $k$ values. The window function used in A15 is constructed in such a way, 
specifically to prevent foregrounds that live at low $k's$ from contaminating higher $k$-modes (see Section \ref{sec:Bias}). 
Minimizing foreground leakage in this way however, comes with the trade-off of compromising power spectrum sensitivity, since 
narrow window functions increases errors for each $k$-mode (\citealt{liu_et_al2014b}).

Determining the source of positive non-EoR detections at higher $k$'s is more difficult. In previous power spectrum results, these detections have been explained as instrumental systematics, particularly time-variable cross talk, RFI, cable reflections, and calibration errors (A15; \citealt{parsons_et_al2014}; \citealt{dillon_et_al2014}; \citealt{beardsley_et_al2016}; \citealt{patil_et_al2017}). In the next section (Section \ref{sec:JackknifeOverview}), we will present some tests that can help distinguish these excesses from that of EoR.

@@

Another type of noise bias can stem from the spurious cross-coupling of signals between antennas. This excess is known as 
instrumental crosstalk and is an inadvertent correlation between two independent measurements via a coupled signal path. 
Crosstalk appears as a constant phase bias in time in visibilities, and it varies slowly compared to the typical fringe-rates of 
sources. Because it is slow-varying, crosstalk can be suppressed using time-averages or fringe-rate filters. However, there 
remains a possibility that power spectrum detections that aren't the cosmological signal are caused by residual, low-level crosstalk which survived any 
suppression techniques.

@@

To mimic the presence of a systematic in part of the measurement, we add a toy sinusoid foreground, similar to the one used in 
Section \ref{sec:toymodel}, to the first $100$ time integrations of both measurements. This represents a foreground signal 
present in, for example, the first half of the LST range used for analysis, but not the second half. Mathematically, if  $\textbf{n}$ 
is noise, $\textbf{e}$ is the EoR signal, and $\textbf{f}$ is the foreground signal, the two measurements (which are cross-multiplied to form power spectra) can be written as:

\begin{equation}
\label{eq:bias1}
\textbf{x}_{a} = \textbf{n}_{a} + \textbf{e} + \textbf{f}
\end{equation}

@@

\begin{equation}
\label{eq:bias2}
\textbf{x}_{b} = \textbf{n}_{b} + \textbf{e} + \textbf{f}.
\end{equation}

@@

\begin{eqnarray}
\textbf{x}_{1} &=& \textbf{n} + \textbf{e} + \textbf{f} \\
\textbf{x}_{2} &=& \textbf{n} + \textbf{e} 
\end{eqnarray}

@@

We form three different power spectrum estimates shown in Figure \ref{fig:toy_bias1}. The first is a null test where we subtract $
\textbf{x}_{2}$ from $\textbf{x}_{1}$ for both measurements $a$ and $b$. This is equivalent to splitting up a full dataset along 
an axis (in this case, time) and subtracting the two to remove sky signal that should ideally be present in both. We cross-multiply measurements $a$ and $b$ to form an un-biased (thermal noise-wise) estimate (blue curve). The second estimate, 
shown in red, is the same null test with the foreground systematic removed (eliminate $\textbf{f}$ in Equations \ref{eq:bias1} 
and \ref{eq:bias2}). Finally, we also show the noise power spectrum (green).

@@

\subsubsection{Signal Loss Methodology} 
\label{sec:siglossmethod}
To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method as in A15, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.\footnote{One 
specific change from A15 is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss, %(by a factor of $\sim$$10$), 
likely due to the use of the fringe-rate filter as a simulator. However, this pipeline difference, while significant, is not the dominant reason why signal loss was under-estimated in A15 (the dominant reason is explained in the main text in Section \ref{sec:Illustration}).}

Suppose that $\textbf{e}$ is a mock injected EoR signal (at some amplitude level), and $\textbf{x}$ is our data. We define $\textbf{r}
$ to be the data plus the EoR signal:
